Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis

### 2024-04-17 paper/paper.qmd
fix some bugs in pdf

```{r dataextract, echo=FALSE, message=FALSE}
# create a table showing the first 20 rows of the cleaned data
income <<- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"))
income %>%
  slice(1:20) %>%
  kable(
    caption = "Extracting the first ten rows from the Income data",
col.names=c("Year", "Geographical location", "Income decile", "Income", "Highest-to-Lowest Average Income Ratio", "income range"),
linesep = "", digits = 1, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)
```

@@

income3 %>%
  filter(
         `Income decile` == "Lowest decile" | `Income decile` == "Total deciles"| `Income decile` == "Highest decile",
         `Geographical_Location` == "Canada") %>%
  ggplot(aes(x=Year, y=Income, colour = `Income decile`))+geom_line()+theme_minimal()+ylab("Average after-tax income")

@@

```{r fig.width=6, fig.height=10}
#| label: fig-planes
#| fig-cap: Analyzing 1976-2021 Income Disparity Trends in Canada by Highest-to-Lowest Average Income Ratio
#| echo: false
#| warning: false
#| message: false
income3 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)
income3 %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical_Location` != "Canada") %>%
  ggplot(aes(x=Year, y=`Highest-to-Lowest_Average_Income_Ratio`, colour = `Geographical_Location`))+geom_point()+theme_minimal()+ylab("Average after-tax income range")
```

@fig-planes, the Highest-to-Lowest Average Income Ratio across Canadian provinces has shown a general upward trend, following a decline from 1980 to 1990. However, each province exhibits distinct performance. Notably, British Columbia's ratio surpasses that of other provinces. Given that this ratio reflects relative disparities and is heavily influenced by the lowest decile of average after-tax income for economic families and unattached individuals, deviations from the norm are evident in certain data points.
```{r}
#| label: fig-planes2
#| fig-cap: Analyzing 1976-2021 Income Disparity Trends in Canada by Average after-tax Income Range
#| echo: false
#| warning: false
#| message: false
income4 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)
income4 %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical_Location` != "Canada") %>%
  ggplot(aes(x=Year, y=`Income_Range`, colour = `Geographical_Location`))+geom_point()+theme_minimal()+ylab("Average after-tax income range")
```

>>>>

```{r dataextract, echo=FALSE, message=FALSE}
# create a table showing the first 20 rows of the cleaned data
income <<- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"))
income %>%
  slice(1:20) %>%
  kable(
    caption = "Extracting the first twenty rows from the Income data",
col.names=c("Year", "Geographical location", "Income decile", "Income", "Highest-to-Lowest Average Income Ratio", "income range"),
linesep = "", digits = 1, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)
```

@@

income3 <- income3 %>% 
  rename(`Geographical Location` = Geographical_Location)

income3 %>%
  filter(
         `Income decile` == "Lowest decile" | `Income decile` == "Total deciles"| `Income decile` == "Highest decile",
         `Geographical Location` == "Canada") %>%
  ggplot(aes(x=Year, y=Income, colour = `Income decile`))+geom_line()+theme_minimal()+ylab("Average after-tax income")

@@

```{r fig.width=6, fig.height=10}
#| label: fig-planes
#| fig-cap: Analyzing 1976-2021 Income Disparity Trends in Canada by Highest-to-Lowest Average Income Ratio
#| echo: false
#| warning: false
#| message: false
income4 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)
income4 <- income4 %>% 
  rename(`Geographical Location` = Geographical_Location)
income4 %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical Location` != "Canada") %>%
  ggplot(aes(x=Year, y=`Highest-to-Lowest_Average_Income_Ratio`, colour = `Geographical Location`))+geom_point()+theme_minimal()+ylab("Average after-tax income range")
```

@fig-planes, the Highest-to-Lowest Average Income Ratio across Canadian provinces has shown a general upward trend, following a decline from 1980 to 1990. However, each province exhibits distinct performance. Notably, British Columbia's ratio surpasses that of other provinces. Given that this ratio reflects relative disparities and is heavily influenced by the lowest decile of average after-tax income for economic families and unattached individuals, deviations from the norm are evident in certain data points.
```{r}
#| label: fig-planes2
#| fig-cap: Analyzing 1976-2021 Income Disparity Trends in Canada by Average after-tax Income Range
#| echo: false
#| warning: false
#| message: false
income5 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)
income5 <- income5 %>% 
  rename(`Geographical Location` = Geographical_Location)
income5 %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical Location` != "Canada") %>%
  ggplot(aes(x=Year, y=`Income_Range`, colour = `Geographical Location`))+geom_point()+theme_minimal()+ylab("Average after-tax income range")
```
